### 2024-05-03 paper/paper.tex
Model feature update
Signed-off-by: Charles Hu <computerscience@verizon.net>

% TODO fill in your paper title
\newcommand{\PaperTitle}{Simple Model Assemblages for Website Identification}
% TODO fill in your paper number when you get it
% \newcommand{\PaperNumber}{XXX}

\documentclass[10pt,sigconf,letterpaper,nonacm]{acmart}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the preamble; include packages as you see fit.
% Here are a few recommendations:
% \usepackage{color}
% \usepackage{graphicx}
% \usepackage[labelformat=simple]{subcaption}
% \usepackage{xspace}
% \usepackage{multirow}
% \usepackage[ruled,vlined]{algorithm2e}
% \usepackage{ulem}
% \normalem

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{\PaperTitle}

\author{Charles Hu}
\email{czhu06@wm.edu}
\affiliation{
  \institution{William \& Mary}
  \city{Williamsburg}
  \state{Virginia}
  \country{USA}
}

\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work. A cite so this compiles~\cite{geneva}.
\end{abstract}

\keywords{Network trafifc, Network traffic classification, Machine learning, Ensemble models}

\maketitle

\section{Introduction}

\section{Proposed Method}

\subsection{Data Specifications}

\subsubsection{Data Collection}

\subsubsection{Data Preprocessing}

\subsubsection{Data Analysis}

\subsection{Model Development}

\subsection{Model Specifications}

\subsubsection{Simple Models}

\subsubsection{Ensemble Models}

\section{Evaluation}

\subsection{Evaluation Metric}

\subsection{Results}

\subsubsection{Baseline}

\subsubsection{Simple Model Results}

\subsubsection{Ensemble Model Results}

\subsubsection{Model Results Overall}

\section{Discussion \& Future Work}

\subsection{Result Interpretations}

\subsection{Future Work}

\section{Conclusion}

% Note from the CFP that this section must include a statement about
% ethical issues; papers that do not include such a statement may be
% rejected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% We're in the endgame now

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}